---
title: "Application Exercise"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: lab.css
link-citations: yes
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE     # for regression output
)
```


1. Open Rstudio Server Pro
2. Create a new project
3. Click File > New File > R Markdown
4. Delete everything except for 
```
---
title: "Untitled"
output: html_document
---
```
5. Give your work a sensible title
6. When you've finished, upload the .html file to Canvas

Generate some "fake" residuals under the following scenarios and plot the Q-Q plot and histogram for each. Describe what you see.

* Left-skewed data
* Right-skewed data
* Data generated under a lognormal distribution (use the `rlnorm()` function)
* Data generated under a t-distribution with 2 degrees of freedom (use the `rt()` function)

Copy the code below to create a design matrix, `X` and an outcome, y. 


```{r}
set.seed(1)
X <- matrix(c(rep(1, 100),
            rnorm(100)), ncol = 2)
y <- 2 + 3 * X[, 2] + 2 * X[, 2]^2 + 0.5 * X[, 2]^3 + rnorm(100)
```


* Add to this design matrix so you can fit a model with a 3rd degree polynomial. 
* Fit this model using linear regression
* Calculate the expected change in Y given a change from the 25th percentile to the 75th percentile
