---
title: "Application Exercise"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: lab.css
link-citations: yes
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE     # for regression output
)
```

_You may use any notes, lectures, or your textbook. Do not work with anyone else on this assessment._

1. Open Rstudio Server Pro
2. Create a new project
3. Click File > New File > R Markdown
4. Delete everything except for 
```
---
title: "Untitled"
output: html_document
---
```
5. Give your work a sensible title
6. When you've finished, upload the .html file to Canvas

Using the `cheddar` data from the faraway package, run the code below to add a time variable. 

```{r}
library(faraway)
cheddar$time <- 1:nrow(cheddar)
```

* Fit a linear model predicting `taste` from `Acetic`, `H2S` and `Lactic`.
* Plot the residuals from the model against `time`. What do you see?
* Fit a GLS model using AR(1) correlation. Is there evidence of correlation between the errors?
