---
title: "Interpretations"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan
</span>
</div> 

---


class: middle

## $\hat\beta$ interpretation in multiple linear regression

--
The coefficient for $x$ is $\hat\beta$ (95% CI: $LB_\hat\beta, UB_\hat\beta$). A one-unit increase in $x$ yields an expected increase in y of $\hat\beta$, **holding all other variables constant**.

---

## Hitters data

```{r, echo = FALSE}
library(ISLR)
summary(lm(Salary ~ AtBat + Hits, data = Hitters))
```

---
class: middle

# standard error

How much we expect the sample slope to vary from one random sample to another.

---

## Hitters data

```{r, echo = FALSE}
library(ISLR)
summary(lm(Salary ~ AtBat + Hits, data = Hitters))
```

---
class: middle

# p-value

The probability of getting a statistic as extreme or more extreme than the observed test statistic **given the null hypothesis is true**

---

## Hitters data

```{r, echo = FALSE}
library(ISLR)
summary(lm(Salary ~ AtBat + Hits, data = Hitters))
```



