---
title: "Indicator Variables"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
  <span>
  Dr. Lucy D'Agostino McGowan
</span>
</div> 

---

```{r, include = FALSE}
library(tidyverse)
```

## Variable types

* So far, our models have only included _numeric_ (_quantitative_) variables
--

  * What would the equation be for predicting $y$ from $x$ when $x$ is numeric?
--

* What would happen if $x$ is categorical?
--

  * What would the equation be for predicting $y$ from $x$ if $x$ is categorical with 2 levels?
--

  * What would the equation be for predicting $y$ from $x$ if $x$ is categorical with 3 levels?
  
---

class: middle

## indicator variable

An **indicator variable** uses two values, usually 0 and 1, to indicate whether a data case does (1) or does not (0) belong to a specific category

---

```{r, echo = TRUE, message = FALSE, warning = FALSE}
library(Stat2Data)
data("Diamonds")
```

.small[
```{r, echo = FALSE}
Diamonds %>% 
  select(TotalPrice, Color, Carat) %>%
  DT::datatable()
```
]

---

## Indicator variables

.question[
What does this line of code do?
]
```{r, echo = TRUE}
Diamonds$ColorD <- ifelse(Diamonds$Color == "D", 1, 0) #<<
Diamonds$ColorE <- ifelse(Diamonds$Color == "E", 1, 0)
Diamonds$ColorF <- ifelse(Diamonds$Color == "F", 1, 0)
Diamonds$ColorG <- ifelse(Diamonds$Color == "G", 1, 0)
Diamonds$ColorH <- ifelse(Diamonds$Color == "H", 1, 0)
Diamonds$ColorI <- ifelse(Diamonds$Color == "I", 1, 0)
Diamonds$ColorJ <- ifelse(Diamonds$Color == "J", 1, 0)
```

---


## Indicator variables

.question[
What does this line of code do?
]
```{r, echo = TRUE}
Diamonds$ColorD <- ifelse(Diamonds$Color == "D", 1, 0) 
Diamonds$ColorE <- ifelse(Diamonds$Color == "E", 1, 0) #<<
Diamonds$ColorF <- ifelse(Diamonds$Color == "F", 1, 0)
Diamonds$ColorG <- ifelse(Diamonds$Color == "G", 1, 0)
Diamonds$ColorH <- ifelse(Diamonds$Color == "H", 1, 0)
Diamonds$ColorI <- ifelse(Diamonds$Color == "I", 1, 0)
Diamonds$ColorJ <- ifelse(Diamonds$Color == "J", 1, 0)
```

---

## Indicator variables

.small[
```{r, echo = FALSE}
Diamonds %>%
  select(TotalPrice, Carat, Color, ColorD, ColorE, ColorF, ColorG, ColorH, ColorI) %>%
  DT::datatable()
```
]

---

## Indicator variables

.question[
What if I wanted to model the relationship between `TotalPrice` and `Color`?
]

.small[
```{r}
Diamonds %>%
  select(TotalPrice, Carat, Color, ColorD, ColorE, ColorF, ColorG, ColorH, ColorI) %>%
  DT::datatable(options = list(pageLength = 5))
```
]

---

## Indicator variables

.question[
Why is `ColorJ` `NA`?
]

.small[
```{r, echo = TRUE, highlight.output = c(9, 10)}
lm(TotalPrice ~ ColorD + ColorE + ColorF + ColorG + ColorH + ColorI + ColorJ,
   data = Diamonds)
```
]

--

* When including indicator variables in a model for `k` categories, always include `k-1` 
--

* The one that is left out is the "reference" category

---

## Indicator variables

.question[
What is the reference category?
]

.small[
```{r, echo = TRUE}
lm(TotalPrice ~ ColorD + ColorE + ColorF + ColorG + ColorH + ColorI,
   data = Diamonds)
```
]

--

* **Interpretation:** A diamond with Color `D` compared to color `J` increases the expected total price by 3632.
--

* **Interpretation:** A diamond with Color `E` compared to color `J` increases the expected total price by 2423

---

## Indicator variables

.question[
What is the reference category?
]

.small[
```{r, echo = TRUE}
lm(TotalPrice ~ ColorD + ColorE + ColorF + ColorG + ColorH + ColorI,
   data = Diamonds)
```
]


* **Interpretation:** A diamond with Color `D` compared to color `J` increases the expected total price by 3632.
* What is the interpretation for a diamond with Color `F`?

---

## In design matrix form

.question[
What would the design matrix look like?
]

$$\begin{bmatrix}1&0&1&0&0&0\\
1&0&0&1&0&0\\
1&0&0&0&0&1\\
1&0&0&1&0&0\\
\vdots&\vdots&\vdots&\vdots&\vdots&\vdots\end{bmatrix}$$

--

.question[
How many columns would this have?
]

--

* The number of indicator variables (-1 for the referent) + 1 for the intercept

---

## R is smart

.small[
```{r, echo = TRUE}
lm(TotalPrice ~ Color, data = Diamonds)
```
]

---

## R is smart

.question[
What is the reference category?
]
.small[
```{r, echo = TRUE}
lm(TotalPrice ~ Color, data = Diamonds)
```
]

--

* What is the interpretation for Color `E` now?

--
* What if we wanted a different referent category?

--
  * We could code the indicators ourselves
  
--
  * We can change the level with the `levels()` function
---

## `r fa("laptop")` `Application Exercise`

Input the **design matrix**, $\mathbf{X}$ from the following data in R

y | age
---|----
2 | "<20"
3 | "20-50"
5 | ">50"
10 | "<20"
1 | "20-50"
3 | ">50"

Estimate the $\beta$ coefficients for predicting y from age.