---
title: "Weighted Least Squares"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan
</span>
</div> 

---

## Weighted Least Squares

* This is a **special case** of GLS where the errors are _uncorrelated_
--

* Now $\Sigma$ is a diagonal matrix with weights $w_i, \dots,w_n$ on the diagonal
--

* $\mathbf{S} = \textrm{diag}(\sqrt{1/w_1}, \dots,\sqrt{1/w_n})$
--

* So we would regress $\sqrt{w_i}y_i$ on $\sqrt{w_i}x_i$
--

* **Note** the column of 1s in the design matrix should now be replaced with $\sqrt{w_i}$
--

* cases with low variability get a high weight and those with high variability a low weight

---

## WLS Examples

* You see a positive relationship when you plot the absolute value of the residuals, $|e_i|$, against a predictor, $x_i$. **Errors are proportional to a predictor** This suggests you should use $w_i = x_i^{-1}$
--

* **Your outcome, $y_i$ is an _average_ of $n_i$ observations**. Here you may want to weight by $w_i = n_i$
--

* **The observed responses are known to be of varying quality.** Here you may want to use $w_i = 1/\textrm{var}(y_i)$

---

## WLS in R

* You can fit weighted least squares by fitting the _normal_ linear model using `lm` and adding the `weights` parameter

```{r, eval = FALSE}
lm(y ~ x, data = data, weights = wts)
```


