---
title: "Building Reports in R Markdown"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan
</span>
</div> 

---

## Harnessing the power of RMarkdown!

* Code chunk options
* Tables
* Inline code
* Output Formats

---

## Code chunk options

* All chunk options can be found on [Yihui's website](http://yihui.name/knitr/options/)
--

* ````{r}`
--

* ````{r, echo = FALSE}`
--

* ````{r, eval = FALSE}`
--

* ````{r, message = FALSE, warning = FALSE}`
--

* ````{r, fig.cap = "PUT CAPTION HERE"}`

---

## Code Chunks

* For these fancy reports, we want to see the R code, but we don't want it interspersed throughout the document. 
* You can "hide" all of your chunks by adding this to the top of your .Rmd file

.small[
````
`r ''````{r, echo = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE)
```
````
]

--

* **THEN** at the end of you document (in the Appendix) you can add

.small[
````
`r ''````{r ref.label=knitr::all_labels(), echo = TRUE, eval = FALSE}
```
````
]
---

class: inverse 
## `r fa("laptop")` `Application Exercise`

* Choose one of your analyses from the last Lab and copy it into a new R Markdown file
* Edit the chunks so that:
  * no messages or warnings are printed
  * All figures have appropriate captions
  * The code all is printed at the end of the document
---

## Tables

* There are several packages for making pretty tables in RMarkdown
--

* [knitr]((https://cran.r-project.org/web/packages/knitr/)
* [xtable](https://cran.r-project.org/web/packages/xtable/)
* [stargazer](https://cran.r-project.org/web/packages/stargazer/)
* [pander](http://rapporter.github.io/pander/)
* [tables](https://cran.r-project.org/web/packages/tables/)
* [ascii](http://eusebe.github.io/ascii/)
---

## Tables with knitr

```{r}
library(knitr)
kable(mtcars[1:4, ], caption = "Look at all those cars!")
```

---

## broom

* Helps create "tidy" tables from your `lm` output

--

.pull-left[
.small[
```{r}
lm(wt ~ am, data = mtcars)
```
]
]

.pull-right[
.small[
```{r}
library(broom)
m <- lm(wt ~ am, data = mtcars)
tidy(m)
```
]

]


---

## kable + broom

* Use the `kable()` function to output a pretty table

.small[
```{r, eval = FALSE}
library(knitr) #<<
m <- lm(wt ~ am, data = mtcars)
kable(tidy(m))
```
]
.small[
```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(knitr) 
m <- lm(wt ~ am, data = mtcars)
kable(tidy(m), format = "html")
```
]
---

class: inverse 

## `r fa("laptop")` `Application Exercise`

* Using the same document as the previous application exercise, update all model output to be in "nicely" formatted tables using `tidy` and `kable` (or a different table package of your choosing)

---

## Inline code

* You can include _inline_ by using `r` 
--

* This is especially useful if you want to reference a numeric value.

---

class: inverse 

## `r fa("laptop")` `Application Exercise`

* Using the same document as the previous application exercise, update all references to numeric values using inline code. _Be sure to round numbers_.

---

## Output formats

## Updating your `yaml`


```{yaml}
---
title: "The title of your document"
name: "Your name"
output: pdf_document #<<
---
```

---

## Output formats

## Updating your `yaml`


```{yaml}
---
title: "The title of your document"
name: "Your name"
output: word_document #<<
---
```

---

class: inverse 

## `r fa("laptop")` `Application Exercise`

* Using the same document as the previous application exercise, update your file to output a word document. 


