---
title: "Visualizing Linear Models"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan
</span>
</div> 

---

## Visualizing linear models

* Forest plots of the coefficients and confidence intervals

--

```{r, echo = FALSE}
library(broom)
library(ggplot2)
mod <- lm(mpg ~ cyl + am + vs + gear, data = mtcars)
mod_tidy <- tidy(mod, conf.int = TRUE)
mod_tidy <- mod_tidy[-1, ] # drop the intercept
mod_tidy$labels <- c("No. cylinders",
                     "Manual transmission",
                     "Straight engine",
                     "No. forward gears")
ggplot(mod_tidy, aes(x = labels, y = estimate, ymin = conf.low, ymax = conf.high, color = ifelse(p.value < 0.05, "p <0.05", "p >0.05"))) +
  geom_pointrange() +
  coord_flip() +
  theme_minimal() +
  scale_color_manual("", 
                     values = c( 
                       "blue", 
                       "black")) + 
  labs(y = "Coefficient (95% CI)", x = "") +
  geom_hline(yintercept = 0, lty = 2) +
  annotate("text", x = 2.25, y = 5.9, label = "Having manual transmission \n(versus automatic) \nyeilds an expected change in \nmiles per gallon of 4.32 \n(95% CI: 0.1, 8.5) holding all other \nvariables constant", size = 2) + 
  annotate("curve", x = 1.5, y = 4.75, xend = 1.2, yend = 4.32,
           arrow = arrow(length = unit(0.03, "npc"))
  )
```

---

## Building up to this plot!

First, fit the model and tidy it

```{r}
library(broom)
library(ggplot2)
mod <- lm(mpg ~ cyl + am + vs + gear, data = mtcars)
mod_tidy <- tidy(mod, conf.int = TRUE)
mod_tidy <- mod_tidy[-1, ] # drop the intercept
```

---

## Building up to this plot!

First, fit the model and tidy it

```{r}
library(broom)
library(ggplot2)
mod <- lm(mpg ~ cyl + am + vs + gear, data = mtcars) #<<
mod_tidy <- tidy(mod, conf.int = TRUE)
mod_tidy <- mod_tidy[-1, ] # drop the intercept
```

---

## Building up to this plot!

First, fit the model and tidy it

```{r}
library(broom)
library(ggplot2)
mod <- lm(mpg ~ cyl + am + vs + gear, data = mtcars)
mod_tidy <- tidy(mod, conf.int = TRUE) #<<
mod_tidy <- mod_tidy[-1, ] # drop the intercept
```

---

## Building up to this plot!

First, fit the model and tidy it

```{r}
library(broom)
library(ggplot2)
mod <- lm(mpg ~ cyl + am + vs + gear, data = mtcars)
mod_tidy <- tidy(mod, conf.int = TRUE)
mod_tidy <- mod_tidy[-1, ] # drop the intercept #<<
```

---

## Create a forest plot


.pull-left[
.small[
```{r, eval = FALSE}
ggplot(mod_tidy,
       aes(x = term, y = estimate, 
           ymin = conf.low,
           ymax = conf.high)) +
  geom_pointrange() #<<
```
]
]

.pull-right[
```{r, echo = FALSE, fig.height = 5}
ggplot(mod_tidy,
       aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 20))
```
]

--
.question[
What could be better?
]

---

## Flip the x- and y-axis

.pull-left[
.small[
```{r, eval = FALSE}
ggplot(mod_tidy,
       aes(x = term, y = estimate, 
           ymin = conf.low,
           ymax = conf.high)) +
  geom_pointrange() +
  coord_flip() #<<
```
]
]

.pull-right[
```{r, echo = FALSE, fig.height = 5}
ggplot(mod_tidy,
       aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  coord_flip() +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 20))
```
]

---

## Axis labels

.pull-left[
.small[
```{r, eval = FALSE}
ggplot(mod_tidy,
       aes(x = term, y = estimate, 
           ymin = conf.low,
           ymax = conf.high)) +
  geom_pointrange() +
  coord_flip() +
  labs(x = "", #<<
       y = "Coefficient (95% CI)") #<<
```
]
]

.pull-right[
```{r, echo = FALSE, fig.height = 5}
ggplot(mod_tidy,
       aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  coord_flip() +
  labs(x = "", y = "Coefficient (95% CI)") +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 20))
```
]

---

## Term labels


.pull-left[
.small[
```{r, eval = FALSE}
mod_tidy$labels <- c(
  "No. cylinders", #<<
  "Manual transmission", #<<
  "Straight engine", #<<
  "No. forward gears") #<<
ggplot(mod_tidy,
       aes(x = labels, y = estimate, #<<
           ymin = conf.low,
           ymax = conf.high)) +
  geom_pointrange() +
  coord_flip() +
  labs(x = "", 
       y = "Coefficient (95% CI)") 
```
]
]

.pull-right[
```{r, echo = FALSE, fig.height = 5}
mod_tidy$labels <- c("No. cylinders",
                     "Manual transmission",
                     "Straight engine",
                     "No. forward gears")
ggplot(mod_tidy,
       aes(x = labels, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  coord_flip() +
  labs(x = "", y = "Coefficient (95% CI)") +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 15))
```
]

--

Better ways to do this (especially if you have lots of variables!)

* `merge()`
* `left_join()`

---

## Add a reference line at the null


.pull-left[
.small[
```{r, eval = FALSE}
ggplot(mod_tidy,
       aes(x = labels, y = estimate, 
           ymin = conf.low,
           ymax = conf.high)) +
  geom_pointrange() +
  geom_hline(yintercept = 0, lty = 2) + #<<
  coord_flip() +
  labs(x = "", 
       y = "Coefficient (95% CI)") 
```
]
]

.pull-right[
```{r, echo = FALSE, fig.height = 5}
ggplot(mod_tidy,
       aes(x = labels, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  coord_flip() +
  labs(x = "", y = "Coefficient (95% CI)") +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 15)) + 
  geom_hline(yintercept = 0, lty = 2)
```
]

---

## Update the theme


.pull-left[
.small[
```{r, eval = FALSE}
ggplot(mod_tidy,
       aes(x = labels, y = estimate, 
           ymin = conf.low,
           ymax = conf.high)) +
  geom_pointrange() +
  geom_hline(yintercept = 0, lty = 2) +
  coord_flip() +
  labs(x = "", 
       y = "Coefficient (95% CI)") +
  theme_minimal() #<<
```
]
]

.pull-right[
```{r, echo = FALSE, fig.height = 5}
ggplot(mod_tidy,
       aes(x = labels, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  coord_flip() +
  geom_hline(yintercept = 0, lty = 2) +
  labs(x = "", y = "Coefficient (95% CI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 15)) 
```
]

--

* Check out the available themes here: https://ggplot2.tidyverse.org/reference/ggtheme.html

---

## Higlight points of interest


.pull-left[
.small[
```{r, eval = FALSE}
ggplot(mod_tidy,
       aes(x = labels, y = estimate, 
           ymin = conf.low,
           ymax = conf.high,
           color = #<<
             ifelse(p.value < 0.05, #<<
                    "p < 0.05", #<<
                    "p > 0.05"))) + #<<
  geom_pointrange() +
  geom_hline(yintercept = 0, lty = 2) +
  coord_flip() +
  scale_color_manual("", #<<
                     values = c( #<<
                       "blue", #<<
                       "black")) + #<<
  labs(x = "", 
       y = "Coefficient (95% CI)") +
  theme_minimal() 
```
]
]

.pull-right[
```{r, echo = FALSE, fig.height = 5}
ggplot(mod_tidy,
       aes(x = labels, y = estimate, ymin = conf.low, ymax = conf.high,          color = 
             ifelse(p.value < 0.05, 
                    "p < 0.05", 
                    "p > 0.05"))) + 
  geom_pointrange() +
  coord_flip() +
  scale_color_manual("", 
                     values = c( 
                       "blue", 
                       "black")) + 
  geom_hline(yintercept = 0, lty = 2) +
  labs(x = "", y = "Coefficient (95% CI)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 15)) 
```
]

---

## Annotate an important point


.pull-left[
.small[
```{r, eval = FALSE}
ggplot(mod_tidy,
       aes(x = labels, y = estimate, 
           ymin = conf.low,
           ymax = conf.high,
           color = 
             ifelse(p.value < 0.05, 
                    "p < 0.05", 
                    "p > 0.05"))) + 
  geom_pointrange() +
  geom_hline(yintercept = 0, lty = 2) +
  coord_flip() +
  scale_color_manual("", 
                     values = c( 
                       "blue", 
                       "black")) + 
  labs(x = "", 
       y = "Coefficient (95% CI)") +
  annotate("text", x = 2.25, y = 5.9, #<<
           label = "Having manual transmission \n(versus automatic) \nyeilds an expected change in \nmiles per gallon of 4.32 \n(95% CI: 0.1, 8.5)", #<<
           size = 2) + #<<
  theme_minimal() 
```
]
]

.pull-right[
```{r, echo = FALSE, fig.height = 5}
ggplot(mod_tidy,
       aes(x = labels, y = estimate, ymin = conf.low, ymax = conf.high,          color = #<<
             ifelse(p.value < 0.05, 
                    "p < 0.05", 
                    "p > 0.05"))) + 
  geom_pointrange() +
  coord_flip() +
  scale_color_manual("", 
                     values = c( 
                       "blue", 
                       "black")) + 
  geom_hline(yintercept = 0, lty = 2) +
  labs(x = "", y = "Coefficient (95% CI)") +
  theme_minimal() +
  annotate("text", x = 1.75, y = 5, 
           label = "Having manual transmission \n(versus automatic) \nyeilds an expected change in \nmiles per gallon of 4.32 \n(95% CI: 0.1, 8.5)\n holding all other \nvariables constant", 
           size = 2) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 15)) 
```
]

---

## Annotate an important point

.pull-left[
.small[
```{r, eval = FALSE}
ggplot(mod_tidy,
       aes(x = labels, y = estimate, 
           ymin = conf.low,
           ymax = conf.high,
           color = 
             ifelse(p.value < 0.05, 
                    "p < 0.05", 
                    "p > 0.05"))) + 
  geom_pointrange() +
  geom_hline(yintercept = 0, lty = 2) +
  coord_flip() +
  scale_color_manual("", 
                     values = c( 
                       "blue", 
                       "black")) + 
  labs(x = "", 
       y = "Coefficient (95% CI)") +
  annotate("text", x = 2.25, y = 5.9,
           label = "Having manual transmission \n(versus automatic) \nyeilds an expected change in \nmiles per gallon of 4.32 \n(95% CI: 0.1, 8.5) holding all other \nvariables constant", 
           size = 2) + 
  annotate("curve", x = 1.5, y = 4.75, #<<
           xend = 1.2, yend = 4.32, #<<
           arrow = arrow(length = unit(0.03, "npc"))) + #<<
  theme_minimal() 
```
]
]

.pull-right[
```{r, echo = FALSE, fig.height = 5}
ggplot(mod_tidy,
       aes(x = labels, y = estimate, ymin = conf.low, ymax = conf.high,          color = #<<
             ifelse(p.value < 0.05, 
                    "p < 0.05", 
                    "p > 0.05"))) + 
  geom_pointrange() +
  coord_flip() +
  scale_color_manual("", 
                     values = c( 
                       "blue", 
                       "black")) + 
  geom_hline(yintercept = 0, lty = 2) +
  labs(x = "", y = "Coefficient (95% CI)") +
  theme_minimal() +
  annotate("text", x = 1.75, y = 5.9, #<<
           label = "Having manual transmission \n(versus automatic) \nyeilds an expected change in \nmiles per gallon of 4.32 \n(95% CI: 0.1, 8.5) holding all other \nvariables constant", 
           size = 2) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 15)) +
  annotate("curve", x = 1.5, y = 4.75, xend = 1.2, yend = 4.32,
           arrow = arrow(length = unit(0.03, "npc")))
```
]

---

## And there you have it!

```{r, echo = FALSE}
ggplot(mod_tidy,
       aes(x = labels, y = estimate, ymin = conf.low, ymax = conf.high,          color = #<<
             ifelse(p.value < 0.05, 
                    "p < 0.05", 
                    "p > 0.05"))) + 
  geom_pointrange() +
  coord_flip() +
  scale_color_manual("", 
                     values = c( 
                       "blue", 
                       "black")) + 
  geom_hline(yintercept = 0, lty = 2) +
  labs(x = "", y = "Coefficient (95% CI)") +
  theme_minimal() +
  annotate("text", x = 2.25, y = 5.9, 
           label = "Having manual transmission \n(versus automatic) \nyeilds an expected change in \nmiles per gallon of 4.32 \n(95% CI: 0.1, 8.5) holding all other \nvariables constant", 
           size = 2) +
  annotate("curve", x = 1.5, y = 4.75, xend = 1.2, yend = 4.32,
           arrow = arrow(length = unit(0.03, "npc")))
```

---

class: inverse

## `r fa("laptop")``Application Exercise`

* Take the model you chose in the previous application exercise and create a figure like the one discussed today. Include a caption that explains what the figure demonstrates.