---
title: "Lab 03 - Multiple Regression Inference"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: lab.css
link-citations: yes
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE     # for regression output
)
```

**Due:** Tuesday 2020-10-13 at 5pm

# Getting started

- Go to RStudio Pro. Create a **New Project** by clicking File > New Project > New Directory. Name the directory something sensible
- Create a new the .Rmd file in this project and you're ready to go!


# Packages

In this lab we will work with one packages: `Stat2Data`.

If you need to install this package, run the following code **once** in your Console:

```{r, eval = FALSE}
install.packages("Stat2Data")
```

If you'd like to run your code in the Console as well you'll also need to load the packages there. To do so, run the following in the console. 

```{r}
library(Stat2Data) 
```

Note that the packages are also loaded with the same commands in your R Markdown document.

# Data

The data frame we will be working with today is called `Backpack` and it is in the `Stat2Data` package. To load the data add the following to your .Rmd file:

```{r}
data("Backpack")
```


To find out more about the dataset, type the following in your Console: `?Backpack`. A question mark before the name of an object will always bring up its help file. This command must be ran in the Console.

```{marginfigure}
Remember: The Console is at the bottom of your RStudio workspace. Things you type in the Console will not be in your final report. This is a good place to peek at data (try typing `Backpack` in the Console) and look at help files with the `?`. 
```

## Conceptual Exercises

1. Describe what a 95% confidence interval means.

2. Find a formula relating $R^2$ and the $F$-test. Show all steps completed using Latex.

3. The hypothesis tests we considered in this class so far require the assumption of _normal errors_. Another way to complete hypothesis testing is via _permutation tests_, which do not require this assumption. To do this:

* Step (1): Fit a your linear model and calculate the test statistic (for example the t-statistic or the F-statistic) for your hypothesis of interest
* Step (2): _Shuffle the responses_ of your outcome variable so they are random and re-fit your linear model using this new outcome. Calculate the test-statistic for your hypothesis of interest for this new model
* Step (3): Repeat step 2 several times and save your test statistic each time
* Step (4): Compare your F-statistic from step (1) to those generated in step (3). Calculate the proportion of test statistics calculated in Step 3 that are as extreme or more extreme than the one calculated in step (1).

Describe the ways in which this method is _similar to_ and _different from_ the bootstrap method we discussed previously.

## Backpack Data

4. Based on the help function, how many rows (n) and how many columns (p) does the `Backpack` data have? What are the variables included in the data frame? Add your responses to your lab report. 

5. Fit a regression model predicting a student's backpack weight from their body weight, year in school, whether they are a graduate or undergraduate student, and the number of credits they are taking that quarter. Use an F-test to determine whether collectively these four predictors have a relationship to the response. Interpret this result.

6. Using the model created in Exercise 5, interpret the coefficient for body weight.

7. Compute 90% and 95% confidence intervals for the predictor body weight. Interpret these.

8. Compute 90% and 95% _bootstrap_ confidence intervals (using the percentile method) for the predictor body weight. How do these compare to those calculated in Exercise 7?

9. Test the null hypothesis that the coefficient for body weight is 0. What is the test statistic? What is the result of this hypothesis test?

10. Fill in the code below to use a **permutation test** to test the null hypothesis that the coefficient for body weight is 0. Interpret this result.

```{r, eval = FALSE}
model <- ## insert your model here from exercise 5
t <- summary(model)$coef["BodyWeight", 3] ## this will pull out the t-stat for BodyWeight

perm_test <- function() {
  m <- lm(sample(BackpackWeight) ~ ---, data = ---) ## fill in with the same model as exercise 5, except sample from BackpackWeight
  return(c(t = summary(m)$coef["BodyWeight", 3]))
}


perms <- purrr::map_df(1:1000, ~ perm_test())

## plot a histogram of the permutations

ggplot(perms, aes(x = t)) +
  geom_histogram(bins = 30) +
  geom_vline(xintercept = c(t, -t))

## calculate the proportion of permutations that are as extreme or more extreme than the one originally calculated
mean(perms$t > t | perms$t < - t)
```

