---
title: "Lab 05: Putting it all together"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: lab.css
link-citations: yes
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE     # for regression output
)
```

**Due:** Tuesday 2020-11-24 at 5pm

# Getting started

- Go to RStudio Pro. Create a **New Project** by clicking File > New Project > New Directory. Name the directory something sensible
- Create a new the .Rmd file in this project and you're ready to go!


# Packages

In this lab we will work with one package: `faraway`.

If you need to install this package, run the following code **once** in your Console:

```{r, eval = FALSE}
install.packages("faraway")
```

If you'd like to run your code in the Console as well you'll also need to load the packages there. To do so, run the following in the console. 

```{r}
library(faraway) 
```

Note that the packages are also loaded with the same commands in your R Markdown document.

# Data

We will work with two data frames today: `stackloss` and `mammalsleep`, both from the`faraway` package. To load the data add the following to your .Rmd file:

```{r}
data("stackloss")
data("mammalsleep")
```

To find out more about the dataset, type the following in your Console: `?stackloss`. A question mark before the name of an object will always bring up its help file. This command must be ran in the Console.

```{marginfigure}
Remember: The Console is at the bottom of your RStudio workspace. Things you type in the Console will not be in your final report. This is a good place to peek at data (try typing `stackloss` in the Console) and look at help files with the `?`. 
```

## Exercises

1. Using all of the tools we've learned so far in this class, use the `stackloss` data set to predict Stack loss using the remaining variables, fitting an appropriate model. Be sure to include:

* A description of the data
* Checks that all of the assumptions of the model you fit are met
* A model that is appropriate for the data at hand (this should include plots with explanations of what they show)
* An explanation for why you chose to use the model chosen
* An interpretation of the final model, including an interpretation of the coefficients for all of the predictors

2. Using all of the tools we've learned so far in this class, use the `mammalsleep` data set to predict sleep using the remaining variables, fitting an appropriate model. Be sure to include:

* A description of the data
* Checks that all of the assumptions of the model you fit are met
* A model that is appropriate for the data at hand (this should include plots with explanations of what they show)
* An explanation for why you chose to use the model chosen
* An interpretation of the final model, including an interpretation of the coefficients for all of the predictors

 
